---
title: "R Notebook"
output: html_notebook
---

## Libraries 

```{r}

library(dplyr)
library(ggplot2)
library(corrplot)
library(RColorBrewer)
library(caTools)

```

## Importing the data

```{r}

data <- read.csv(file = "CCHS_Annual.csv")

```

## Removing non-relevant columns 

```{r}

data<- subset(data,select = c(DHH_SEX,DHHGAGE,VERDATE, HWTDGISW,GEN_015,GEN_020, INCDGHH, WDM_005, WDM_010, WDM_015, WDM_030, PAADVACV, EHG2DVR3, ALCDVTTM,DHHDGHSZ, DHHGMS,LBFDVWSS, SMKDVSTY,SDCDVIMM,PAY_045, GEO_PRV, GEODGHR4, ADM_045, CCC_140, CCC_095, CCC_065, CIH_030H, DOFSC, CCC_085, CCC_130, DOCP2, UPE_010, SDCDGCB, FVCDVFRU, PHC_040,GEN_005))
data

```

# Structure of R dataframe

```{r}

str(data)

```


```{r , echo=FALSE}

data <- data  %>% 
  mutate(HWTDGISW = case_when(data$HWTDGISW >=  'Normal weight' ~ 1,
                                  data$HWTDGISW >=  'Obese - Class I, II, III'~ 2,
                                  data$HWTDGISW >=  'Overweight '~ 3,
                                  data$HWTDGISW >= 'Underweight'~ 4))

str(data)

```

```{r , echo=FALSE}
data <- data  %>% 
  mutate(SMKDVSTY = case_when(data$SMKDVSTY >=  'Current daily smoker' ~ 1,
                                  data$SMKDVSTY >=  'Current occasional smoker'~ 2,
                                  data$SMKDVSTY >=  'Experimental smoker (at least 1 cig, non-smoker now)'~ 3,
                                  data$SMKDVSTY >= 'Former daily smoker (non-smoker now)l'~ 4,
                                  data$SMKDVSTY >= 'Former occasional smoker (non-smoker now)'~ 5,
                                  data$SMKDVSTY <= 'Lifetime abstainer (never smoked a whole cigarette)'~ 6))
str(data)

```

```{r , echo=FALSE}
data <- data  %>% 
  mutate(GEN_020 = case_when(data$GEN_020 >=  'A bit stressful' ~ 1,
                                  data$GEN_020 >=  'Extremely stressful'~ 2,
                                  data$GEN_020 >=  'Not at all stressful '~ 3,
                                  data$GEN_020 >= 'Not very stressful'~ 4,
                                  data$GEN_020 >= 'Not very stressful'~ 5,
                                  data$GEN_020 <= 'Quite a bit stressful'~ 6))

str(data)

```

```{r , echo=FALSE}
data <- data  %>% 
  mutate(ALCDVTTM = case_when(data$ALCDVTTM >=  'Did not drink in the last 12 months' ~ 1,
                                  data$ALCDVTTM >=  'Occasional drinker'~ 2,
                                  data$ALCDVTTM >=  'Regular drinker'~ 3))

str(data)

```

```{r , echo=FALSE}

data <- data  %>% 
  mutate(INCDGHH = case_when(data$INCDGHH >=  '$20,000 to $39,999' ~ 1,
                                  data$INCDGHH >=  '$40,000 to $59,999'~ 2,
                                  data$INCDGHH >=  '$60,000 to $79,999 '~ 3,
                                  data$INCDGHH >= '$80,000 or more'~ 4,
                                  data$INCDGHH >= 'No income or less than $20,000'~ 5))

str(data)

```


```{r , echo=FALSE}

unique(data$GEN_015)
data <- data  %>% 
  mutate(GEN_015 = case_when(data$GEN_015 >=  '0' ~ 0,
                                  data$GEN_015 >=  'Poor'~ 1,
                                  data$GEN_015 >=  'Fair'~ 2,
                                  data$GEN_015 >=  'Good'~ 3,
                                  data$GEN_015 >=  'Very good'~ 4,
                                  data$GEN_015 >=  'Excellent'~ 5))

str(data)

```
```{r , echo=FALSE}

data <- data  %>% 
  mutate(WDM_005 = case_when(data$WDM_005 >=  'A lot of difficulty' ~ 1,
                                  data$WDM_005 >=  'Cannot do at all / Unable to do' ~2,
                                  data$WDM_005 >=  'No difficulty '~ 3,
                                  data$WDM_005 >= 'Some difficulty'~ 4))

data <- data  %>% 
  mutate(WDM_010 = case_when(data$WDM_010 >=  'A lot of difficulty' ~ 1,
                                  data$WDM_010 >=  'Cannot do at all / Unable to do' ~2,
                                  data$WDM_010 >=  'No difficulty '~ 3,
                                  data$WDM_010 >= 'Some difficulty'~ 4))

data <- data  %>% 
  mutate(WDM_015 = case_when(data$WDM_015 >=  'A lot of difficulty' ~ 1,
                                  data$WDM_015 >=  'Cannot do at all / Unable to do' ~2,
                                  data$WDM_015 >=  'No difficulty '~ 3,
                                  data$WDM_015 >= 'Some difficulty'~ 4))

data <- data  %>% 
  mutate(WDM_030  = case_when(data$WDM_030  >=  'A lot of difficulty' ~ 1,
                                  data$WDM_030  >=  'Cannot do at all / Unable to do' ~2,
                                  data$WDM_030  >=  'No difficulty '~ 3,
                                  data$WDM_030  >= 'Some difficulty'~ 4))

str(data)

```
```{r , echo=FALSE}
data <- data  %>% 
  mutate(PAADVACV = case_when(data$PAADVACV >=  'No physical activity minutes reported' ~ 1,
                                  data$PAADVACV >=  'Physically active at / above recommended level from CPAG'~ 2,
                                  data$PAADVACV >=  'Physically active below recommended level from CPAG'~ 3))

str(data)

```
```{r , echo=FALSE}
data <- data  %>% 
  mutate(EHG2DVR3 = case_when(data$EHG2DVR3 >=  'Less than secondary school graduation' ~ 1,
                                  data$EHG2DVR3 >=  'Post-secondary certificate diploma or univ degree'~ 2,
                                  data$EHG2DVR3 >=  'Secondary school graduation, no post-secondary education'~ 3))

str(data)

```

```{r , echo=FALSE}
data <- data  %>% 
  mutate(DHHGMS = case_when(data$DHHGMS >=  'Common-law' ~ 1,
                                  data$DHHGMS >=  'Married'~ 2,
                                  data$DHHGMS >=  'Single'~ 3,
                                  data$DHHGMS >=  'Widowed/Divorced/Separated'~ 4))

str(data)

```


```{r , echo=FALSE}
data <- data  %>% 
  mutate(LBFDVWSS = case_when(data$LBFDVWSS >=  '0' ~ 1,
                                  data$LBFDVWSS >=  'Did not have a job - last week'~ 2,
                                  data$LBFDVWSS >=  'Worked at a job / business - last week'~ 3))

str(data)

```


```{r , echo=FALSE}
data <- data  %>% 
  mutate(SDCDVIMM = case_when(data$SDCDVIMM >=  'Landed immigrant / non-permanent resident' ~ 1,
                                  data$SDCDVIMM >=  'Non-immigrant (Canadian born)'~ 2))

str(data)

```

```{r , echo=FALSE}

data <- data  %>% 
  mutate(CCC_095  = case_when(data$CCC_095  >=  'Yes' ~ 1,
                                  data$CCC_095  >=  'No'~ 2))
data <- data  %>% 
  mutate(CCC_065   = case_when(data$CCC_065   >=  'Yes' ~ 1,
                                  data$CCC_065   >=  'No'~ 2))

data <- data  %>% 
  mutate(DOFSC  = case_when(data$DOFSC  >=  'Yes' ~ 1,
                                  data$DOFSC  >=  'No'~ 2))
data <- data  %>% 
  mutate(CCC_085   = case_when(data$CCC_085   >=  'Yes' ~ 1,
                                  data$CCC_085   >=  'No'~ 2))
data <- data  %>% 
  mutate(CCC_130   = case_when(data$CCC_130   >=  'Yes' ~ 1,
                                  data$CCC_130   >=  'No'~ 2))
data <- data  %>% 
  mutate(DOCP2     = case_when(data$DOCP2     >=  'Yes' ~ 1,
                                  data$DOCP2     >=  'No'~ 2))

data <- data  %>% 
  mutate(SDCDGCB  = case_when(data$SDCDGCB   >=  'Canada' ~ 1,
                                  data$SDCDGCB >=  'Other'~ 2))
str(data)

```

```{r , echo=FALSE}
data <- data  %>% 
  mutate(PHC_040 = case_when(data$PHC_040 >=  'Another language' ~ 1,
                                  data$PHC_040 >=  'English'~ 2,
                                  data$PHC_040 >=  'English and another language)'~ 3,
                                  data$PHC_040 >= 'English and French'~ 4,
                                  data$PHC_040 >= 'French'~ 5,
                                  data$PHC_040 <= 'French and another language)'~ 6))
str(data)

```
## Missing Value Treatment

```{r}
is.na(data)<-sapply(data, is.infinite)
data[is.na(data)]<-0
sum(is.na(data))
mean(is.na(data))

```


```{r}

data <- na.omit(data)
sum(is.na(data))

```


# 1. Mental Health of Immigrants in Canada:

● It will analyze how lifestyle has changed for immigrants and how it
affects their mental health, other habits such as smoking, drinking, and
on their body such as their BMI. How often they contact health
professionals, household income, food security, Immigrants status,
how long since moved, stress, mood disorder

● Also we can consider comparing the ratio of smoking, drinking,stress of
Immigrants to Canadian Residents across age and gender.

```{r}

anov <- aov(HWTDGISW ~ (GEN_020 + SMKDVSTY + ALCDVTTM), data = data, na.action=na.exclude)
summary(anov)

```


```{r}

ggplot(data,aes(y=SMKDVSTY+ALCDVTTM, x=SDCDVIMM ,color=factor(DHHGAGE)))+geom_point()+stat_smooth(method="lm",se=FALSE)

```

```{r}

ggplot(data,aes(y=SMKDVSTY+ALCDVTTM, x=SDCDVIMM,color=factor(DHH_SEX )))+geom_point()+stat_smooth(method="lm",se=FALSE)

```
# 2. Diseases
● Analyzing the ratio of selected health conditions such as Diabetes,
Health Diseases, Cancer, Blood Pressure among Immigrants vs
Canadian Residents for Men and Women above 21.


```{r}

data$age  <- data$DHHGAGE
data$age  <- ifelse(data$age  %in% c("Age between 12 and 14","Age between 18 and 19", "Age between 15 and 17"), 1, 2)

subset<- data[data$age == 2, ] 

```
```{r}

ggplot(subset,aes(y=CCC_095 + CCC_085 + CCC_130+ CCC_065, x=SDCDVIMM,color=factor(DHH_SEX )))+geom_point()+stat_smooth(method="lm",se=FALSE)

```

## Distribution of few characteristics for Men and Women across different age groups, different geographic locations.
● Physical Activity: Active, Moderate, Inactive
● Contact Health Professionals: Received health care in 12 months ,Didn’t receive health care
● Neurological Condition:
● Functional Health: Moderate, Severe, Good, Perfect

```{r}
data$GEODGHR4 <- as.numeric(as.factor(data$GEODGHR4))

ggplot(data,aes(y=age+ GEODGHR4, x=PAADVACV + GEN_015 + GEN_015 + DOCP2,color=factor(DHH_SEX ))) +geom_point()+stat_smooth(method="lm",se=FALSE)

```
## In terms of Healthy Lifestyle:
● Compared on basis of sex across different age group, different
geographic location
● Components that can be taken into consideration:
Citizenship of Person, BMI, Biking and Helmet for Biking, Fruit dietary,
Leisure time Physical Activity, level of form

```{r}
data$UPE_010 <- as.numeric(as.factor(data$UPE_010))
data$PAY_045 <- as.numeric(as.factor(data$PAY_045))
ggplot(data,aes(y=age+ GEODGHR4,x=HWTDGISW+ UPE_010+ SDCDGCB+ PAY_045+FVCDVFRU ,color=factor(DHH_SEX)))+geom_point()+stat_smooth(method="lm",se=FALSE)

```
## Chronic Diseases for Aboriginal Population:
● Based on visiting doctor or not, receiving service or not, income, age
```{r}
ggplot(data,aes(y=CCC_095 + CCC_085 + CCC_130+ CCC_065 ,x=age+ DOCP2+ INCDGHH ,color=factor(DHHGAGE)))+geom_point()+stat_smooth(method="lm",se=FALSE)

```

# Distribution of selected characteristic by immigration status, household population for above 18.
● Depression
● Alcohol Consumption
● Immigration Status
● Length of Immigration
● Age
● Birth Region
● Sex
● Marital Status
● Education
● Household Income
● Work Status
● Official Language
● Neurological

```{r}

data$age  <- data$DHHGAGE
data$age  <- ifelse(data$age  %in% c("Age between 12 and 14","Age between 15 and 17"), 1, 2)
data$DHHDGHSZ <- as.numeric(as.factor(data$DHHDGHSZ))

str(data)
```


```{r}
ggplot(data,aes(y=SDCDVIMM+ DHHDGHSZ,x=PHC_040 + HWTDGISW + PAADVACV+ ALCDVTTM + INCDGHH +GEN_015 + SDCDVIMM + SDCDGCB+LBFDVWSS + DHHGMS  ,color=factor(DHHGAGE)))+geom_point()+stat_smooth(method="lm",se=FALSE)
```




